---
title: "P8105 HW6"
author: "Chee Kay Cheong"
date: "2022-12-03"
output: github_document
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)
library(modelr)
library(mgcv)
```

## Problem 2

Load and clean the `homicide` dataset:
```{r load and clean homicide dataset}
homicide = 
  read_csv("./Data/homicide_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    across(c(victim_age, victim_sex, victim_race), na_if, "Unknown")) %>% 
  mutate(
    victim_age = as.numeric(victim_age),
    reported_date = as.character(reported_date),
    reported_date = as.Date(reported_date, format = "%Y%m%d"),
    state = str_replace(state, "w", "W"),
    case_solved = ifelse(disposition == "Closed by arrest", 1, 0)) %>%
  unite(col = 'city_state', c('city', 'state'), sep = ', ') %>% 
  subset(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")) %>% 
  subset(victim_race %in% c("White", "Black"))
```

##### Baltimore, MD

```{r create baltimore dataset}
baltimore = 
  homicide %>% 
  filter(
    city_state == "Baltimore, MD") %>%
  mutate(
    victim_race = fct_relevel(victim_race, "White")) %>% 
  select(case_solved, victim_age, victim_sex, victim_race)
```

```{r logistic regression for baltimore}
logreg_balti = 
  baltimore %>% 
  glm(case_solved ~ victim_age + victim_sex + victim_race, data = . , family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error))

logreg_balti %>% 
  select(-std.error, -statistic) %>% 
  knitr::kable(digits = 2)
```

##### All cities

```{r all cities summary}
all_cities = 
  homicide %>% 
    nest(df = -city_state) %>% 
    mutate(
    # Create a new variable that map the logistic regression function to each of the cities.
      models = map(.x = df, ~glm(case_solved ~ victim_age + victim_sex + victim_race, data = . , family = binomial())),
    # Then, create another variable that shows the tidy results of the logistic regression.
      results = map(models, broom::tidy)) %>% 
  select(city_state, results) %>% 
  unnest(results) %>%
  filter(term == "victim_sexMale") %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error))
```

```{r plot OR AND CIs, fig.width = 8, fig.height = 6}
all_cities %>% 
  select(city_state, term, OR, CI_lower, CI_upper) %>% 
  mutate(
    city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = OR, y = city_state)) + 
    geom_vline(aes(xintercept = 1), size = 0.25, linetype = "dashed") + 
    geom_errorbar(aes(xmax = CI_upper, xmin = CI_lower), size = 0.5, height = 0.2, color = "gray50") +
    geom_point(size = 1.5, color = "orange") +
    theme_bw()+
    theme(panel.grid.minor = element_blank()) +
    ylab("City, State") +
    xlab("Odds ratio") +
    ggtitle("Odds ratio comparing homicide resolved among male to female in each city")
  
```


## Problem 3

Load and clean the `birthweight` dataset:
```{r}
bwt_df = 
  read_csv("./Data/birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    babysex = ifelse(babysex == 1, "Male", "Female"),
    babysex = as.factor(babysex),
    malform = ifelse(malform == 1, "present", "absent"),
    malform = as.factor(malform),
    delwt = delwt * 453.592) # I want to convert pounds to grams

# I did not change anything for father's & mother's race because I don't plan to use them in my regression model.
```

##### Proposed regression model for birthweight

Outcome of interest: 
`bwt` (baby's birthweight - grams)

Predictors of interest:                                                                                                                      
* `delwt` (mother's weight at delivery - grams)                                                                                             
* `gaweeks` (gestational age in weeks)                                                                                                       
* `momage` (mother’s age at delivery - years)

Because the outcome of interest and all the predictors of interest are continuous variables, I plan to use linear regression
to determine the the relationship between the predictors of interest and baby birthweight.

Proposed linear regression model:

Birthweight = intercept + beta1(`delwt`) + beta2(`gaweeks`) + beta3(`momage`)

```{r create hypothetical dataset}
hypo_df = 
  bwt_df %>% 
  select(bwt, delwt, gaweeks, momage)

linear_model = 
  hypo_df %>% 
  lm(bwt ~ delwt + gaweeks + momage, data = .)

hypo_resid = 
  hypo_df %>% 
  mutate(
  add_residuals(hypo_df, linear_model),
  add_predictions(hypo_df, linear_model))
```

```{r residual plot}
hypo_resid %>% 
ggplot(aes(x = pred, y = resid)) +
  geom_point() +
  labs(
    title = 'Residual vs. Predicted Values Plot', 
    x = 'Predicted Values', 
    y = 'Residuals')
```




